---
title: "Comparing Best RF Models"
author: "Kirsten Dimmler"
format:
  html:
    toc: true
    number-sections: true
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(patchwork)
library(RColorBrewer)
library(ComplexUpset)
```

# Top Model

## Training Summary

```{r}
train <- read.csv('RF_summaries/Combined_TopModel_Training.csv', header=TRUE)

train_scores <- train |> select(run, cv_f1)

train_scores_unique <- train_scores %>%
  distinct(run, .keep_all = TRUE)

train <- train %>%
  select(-cv_f1, -run) %>%
  mutate(
    across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(train, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# Summarize F1 stats per run
train_summary <- train_scores %>%
  group_by(run) %>%
  summarise(
    mean_f1 = mean(cv_f1, na.rm = TRUE),
    min_f1 = min(cv_f1, na.rm = TRUE),
    max_f1 = max(cv_f1, na.rm = TRUE)
  )

plot_f1 <- ggplot(train_scores_unique, aes(x = factor(run), y = cv_f1)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "F1 Score", title = "F1 Scores") +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Training - Top Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Test Summary

```{r}
test <- read.csv('RF_summaries/Combined_TopModel_Test.csv', header=TRUE)

test_scores <- test |> select(run, test_f1)

test_scores_unique <- test_scores %>%
  distinct(run, .keep_all = TRUE)

test <- test %>%
  select(-test_f1, -run) %>%
  mutate(across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(test, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

plot_f1 <- ggplot(test_scores_unique, aes(x = factor(run), y = test_f1)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "F1 Score", title = "F1 Scores") +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Test - Top Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Validation Summary

```{r}
val <- read.csv('RF_summaries/Combined_TopModel_Validation.csv', header=TRUE)

val_scores <- val |> select(run, val_f1)

val_scores_unique <- val_scores %>%
  distinct(run, .keep_all = TRUE)

val <- val %>%
  select(-val_f1, -run) %>%
  mutate(across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(val, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

plot_f1 <- ggplot(val_scores_unique, aes(x = factor(run), y = val_f1)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "F1 Score", title = "F1 Scores") +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Validation - Top Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## All Datasets Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(train, test, val))

all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(50, 100, 500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(3, 7, 12, 20, 50, 100)),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 50, 100, 200, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20, 50))
  )


# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: F1 Score Plot ---
combined_scores <- train_scores %>%
  left_join(test_scores, by = "run") %>%
  left_join(val_scores, by = "run")

scores_long <- combined_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean F1 scores
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(mean_f1 = mean(f1, na.rm = TRUE), .groups = "drop")

# Plot with text labels for mean
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    color = 'black',
    stat = "identity"
  ) +
  geom_text(
    data = summary_scores,
    aes(
      x = factor(run),
      y = mean_f1 + 0.04,
      label = sprintf("%.2f", mean_f1),
      group = set
    ),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_f1" = "#1f77b4", "test_f1" = "#2ca02c", "val_f1" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "F1 Scores (Mean Highlighted)",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

final_plot


```

## Validation and Test Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

# Convert to factors with correct level ordering
all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(50, 100, 500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(3, 7, 12, 20, 50, 100)),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 50, 100, 200, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20, 50))
  )


# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: F1 Score Plot ---
combined_scores <- test_scores_unique %>%
  left_join(val_scores_unique, by = "run")

scores_long <- combined_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean F1 scores
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(mean_f1 = mean(f1, na.rm = TRUE), .groups = "drop")

# Plot with text labels for mean
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    color = 'black',
    stat = "identity"
  ) +
  geom_text(
    data = summary_scores,
    aes(
      x = factor(run),
      y = mean_f1 + 0.04,
      label = sprintf("%.2f", mean_f1),
      group = set
    ),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_f1" = "#1f77b4", "test_f1" = "#2ca02c", "val_f1" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "F1 Scores (Mean Highlighted)",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "Test and Validation Datasets – Top Models",
  )

final_plot
```

# Top 5 Models

## Training Summary

```{r}
train <- read.csv('RF_summaries/Combined_Top5_Training.csv', header=TRUE)

train_scores <- train |> select(run, cv_f1)

# train_scores_unique <- train_scores %>%
#   distinct(run, .keep_all = TRUE)

train <- train %>%
  select(-cv_f1, -run) %>%
  mutate(
    # Recode first, then convert to factor
    max_features = case_when(
      max_features == "1" ~ "0.5",
      max_features == "2" ~ "sqrt",
      max_features == "3" ~ "log2",
      TRUE ~ as.character(max_features)  # <-- required fallback
    ),
    criterion = case_when(
      criterion == "1" ~ "gini",
      criterion == "2" ~ "entropy",
      TRUE ~ as.character(criterion)
    )
  ) %>%
  mutate(across(everything(), as.factor)) 
```

```{r}
#| fig-height: 8
#| fig-width: 8

train_scores_grouped <- train_scores %>%
  group_by(run) %>%
  mutate(instance = row_number()) %>%
  ungroup()

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(train, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

train_summary <- train_scores_grouped %>%
  group_by(run) %>%
  summarise(
    min_f1 = min(cv_f1, na.rm = TRUE),
    max_f1 = max(cv_f1, na.rm = TRUE),
    mean_f1 = mean(cv_f1, na.rm = TRUE),
    .groups = "drop"
  )

plot_f1 <- ggplot(train_summary, aes(x = factor(run), y = mean_f1)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_errorbar(
    aes(ymin = min_f1, ymax = max_f1),
    width = 0.2,
    color = "black",
    size = 0.6
  ) +
  geom_text(
    aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
    size = 3.5
  ) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "F1 Score per Run (Mean with Range)"
  ) +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Training - Top 5 Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Test Summary

```{r}
test <- read.csv('RF_summaries/Combined_Top5_Test.csv', header=TRUE)

test_scores <- test |> select(run, test_f1)

test_scores_unique <- test_scores %>%
  distinct(run, .keep_all = TRUE)

test <- test %>%
  select(-test_f1, -run) %>%
  mutate(
    # Recode first, then convert to factor
    max_features = case_when(
      max_features == "1" ~ "0.5",
      max_features == "2" ~ "sqrt",
      max_features == "3" ~ "log2",
      TRUE ~ as.character(max_features)  # <-- required fallback
    ),
    criterion = case_when(
      criterion == "1" ~ "gini",
      criterion == "2" ~ "entropy",
      TRUE ~ as.character(criterion)
    )
  ) %>%
  mutate(across(everything(), as.factor)) 
```

```{r}
#| fig-height: 8
#| fig-width: 8

test_scores_grouped <- test_scores %>%
  group_by(run) %>%
  mutate(instance = row_number()) %>%
  ungroup()

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(test, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

test_summary <- test_scores_grouped %>%
  group_by(run) %>%
  summarise(
    min_f1 = min(test_f1, na.rm = TRUE),
    max_f1 = max(test_f1, na.rm = TRUE),
    mean_f1 = mean(test_f1, na.rm = TRUE),
    .groups = "drop"
  )

plot_f1 <- ggplot(test_summary, aes(x = factor(run), y = mean_f1)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_errorbar(
    aes(ymin = min_f1, ymax = max_f1),
    width = 0.2,
    color = "black",
    size = 0.6
  ) +
  geom_text(
    aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
    size = 3.5
  ) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "F1 Score per Run (Mean with Range)"
  ) +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Test - Top 5 Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Validation Summary

```{r}
val <- read.csv('RF_summaries/Combined_Top5_Validation.csv', header=TRUE)

val_scores <- val |> select(run, val_f1)

val_scores_unique <- val_scores %>%
  distinct(run, .keep_all = TRUE)

val <- val %>%
  select(-val_f1, -run) %>%
  mutate(
    # Recode first, then convert to factor
    max_features = case_when(
      max_features == "1" ~ "0.5",
      max_features == "2" ~ "sqrt",
      max_features == "3" ~ "log2",
      TRUE ~ as.character(max_features)  # <-- required fallback
    ),
    criterion = case_when(
      criterion == "1" ~ "gini",
      criterion == "2" ~ "entropy",
      TRUE ~ as.character(criterion)
    )
  ) %>%
  mutate(across(everything(), as.factor)) 
```

```{r}
#| fig-height: 8
#| fig-width: 8

val_scores_grouped <- val_scores %>%
  group_by(run) %>%
  mutate(instance = row_number()) %>%
  ungroup()

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(val, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

val_summary <- val_scores_grouped %>%
  group_by(run) %>%
  summarise(
    min_f1 = min(val_f1, na.rm = TRUE),
    max_f1 = max(val_f1, na.rm = TRUE),
    mean_f1 = mean(val_f1, na.rm = TRUE),
    .groups = "drop"
  )

plot_f1 <- ggplot(val_summary, aes(x = factor(run), y = mean_f1)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_errorbar(
    aes(ymin = min_f1, ymax = max_f1),
    width = 0.2,
    color = "black",
    size = 0.6
  ) +
  geom_text(
    aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
    size = 3.5
  ) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "F1 Score per Run (Mean with Range)"
  ) +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Validation - Top 5 Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## All Datasets Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(train, test, val))

all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(50, 100, 500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(3, 7, 12, 20, 50, 100)),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 50, 100, 200, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20, 50))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: F1 Score Plot ---
# combined_scores <- train_scores_unique %>%
#   left_join(test_scores_unique, by = "run") %>%
#   left_join(val_scores_unique, by = "run")

combined_scores <- train_scores %>%
  left_join(test_scores, by = "run") %>%
  left_join(val_scores, by = "run")

scores_long <- combined_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean, min, and max F1 scores
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(
    mean_f1 = mean(f1, na.rm = TRUE),
    ymin = min(f1, na.rm = TRUE),
    ymax = max(f1, na.rm = TRUE),
    .groups = "drop"
  )

# Final plot with bars, error bars, and mean labels
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    # color = 'black',
    stat = "identity"
  ) +
  geom_errorbar(
    data = summary_scores,
    aes(x = factor(run), ymin = ymin, ymax = ymax, group = set),
    position = position_dodge(width = 0.7),
    width = 0.25,
    size = 0.75,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text(
    data = summary_scores,
    aes(x = factor(run), y = ymax + 0.06, label = sprintf("%.2f", mean_f1), group = set),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_f1" = "#1f77b4", "test_f1" = "#2ca02c", "val_f1" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "Mean F1 Scores and Range",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

final_plot
```

## Validation and Test Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(50, 100, 500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(3, 7, 12, 20, 50, 100)),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 50, 100, 200, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20, 50))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: F1 Score Plot ---
combined_scores <- test_scores %>%
  left_join(val_scores, by = "run")

scores_long <- combined_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean, min, and max F1 scores
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(
    mean_f1 = mean(f1, na.rm = TRUE),
    ymin = min(f1, na.rm = TRUE),
    ymax = max(f1, na.rm = TRUE),
    .groups = "drop"
  )

# Final plot with bars, error bars, and mean labels
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    # color = 'black',
    stat = "identity"
  ) +
  geom_errorbar(
    data = summary_scores,
    aes(x = factor(run), ymin = ymin, ymax = ymax, group = set),
    position = position_dodge(width = 0.7),
    width = 0.25,
    size = 0.75,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text(
    data = summary_scores,
    aes(x = factor(run), y = ymax + 0.06, label = sprintf("%.2f", mean_f1), group = set),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_f1" = "#1f77b4", "test_f1" = "#2ca02c", "val_f1" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "F1 Score",
    title = "Mean F1 Scores and Range",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

final_plot
```

## Upset Plot

```{r}
#| fig-height: 8
#| fig-width: 8

# Step 1: Filter and add a unique row ID for each model
filtered <- all %>%
  select(
    n_estimators,
    max_depth,
    min_samples_split,
    min_samples_leaf,
    criterion,
    max_features
  ) %>%
  filter(
    criterion == "entropy",
    max_features == 0.5
  ) %>%
  select(-criterion, -max_features) %>%
  mutate(row_id = row_number())  # Add unique row ID

# Step 2: Convert to long format and flag presence
filtered_long <- filtered %>%
  pivot_longer(cols = -row_id, names_to = "param", values_to = "value") %>%
  mutate(present = TRUE) %>%
  unite("feature", param, value, sep = "_")

# Step 3: Convert to binary wide format
binary_data <- filtered_long %>%
  pivot_wider(
    id_cols = row_id,
    names_from = feature,
    values_from = present,
    values_fill = list(present = FALSE)
  ) %>%
  select(-row_id)

# Step 4: Make the upset plot
upset(
  binary_data,
  intersect = colnames(binary_data),
  name = "Hyperparameter Combination",
  height_ratio = 1.3
)
```

```{r}
#| fig-height: 8
#| fig-width: 8

# Step 1: Filter and add a unique row ID for each model
filtered <- all %>%
  select(
    n_estimators,
    max_depth,
    min_samples_split,
    min_samples_leaf,
    criterion,
    max_features
  ) %>%
  filter(
    criterion == "entropy",
    max_features == 0.5
  ) %>%
  select(-criterion, -max_features) %>%
  mutate(row_id = row_number())  # Add unique row ID

# Step 2: Convert to long format and flag presence
filtered_long <- filtered %>%
  pivot_longer(cols = -row_id, names_to = "param", values_to = "value") %>%
  mutate(present = TRUE) %>%
  unite("feature", param, value, sep = "_")

# Step 3: Convert to binary wide format
binary_data <- filtered_long %>%
  pivot_wider(
    id_cols = row_id,
    names_from = feature,
    values_from = present,
    values_fill = list(present = FALSE)
  ) %>%
  select(-row_id)

# Step 4: Make the upset plot
upset(
  binary_data,
  intersect = colnames(binary_data),
  name = "Hyperparameter Combination",
  height_ratio = 1.3,
  min_size = 2
)
```

I chose the right-most combination because it contains more hyperparameter values that were chosen more often by good models.
