---
title: "Comparing Best RF Models"
author: "Kirsten Dimmler"
format:
  html:
    toc: true
    number-sections: true
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(patchwork)
library(RColorBrewer)
library(ComplexUpset)
library(gt)
```

# Top Model

## Training Summary

```{r}
cc <- c(max_features = "character")  # add more like criterion="character" if needed

read_one <- function(path) {
  read.csv(path, header = TRUE, stringsAsFactors = FALSE, colClasses = cc)
}

train1 <- read_one('final_plots/RF1_Top5_CV.csv')
train2 <- read_one('../tuning2/final_plots/RF1_Top5_CV.csv')
train3 <- read_one('../tuning3/final_plots/RF1_Top5_CV.csv')
train4 <- read_one('../tuning4/final_plots/RF1_Top5_CV.csv')
train5 <- read_one('../tuning5/final_plots/RF1_Top5_CV.csv')

train1$run <- 'run1'
train2$run <- 'run2'
train3$run <- 'run3'
train4$run <- 'run4'
train5$run <- 'run5'

train <- dplyr::bind_rows(train1, train2, train3, train4, train5)

train <- train %>%
  mutate(
    max_depth = factor(max_depth,
                       levels = 1:5,
                       labels = c("10", "12", "16", "20", "None")),
    max_features = factor(max_features,
                          levels = 1:4,
                          labels = c("0.3", "0.5", "sqrt", "log2")),
    criterion = factor(criterion,
                       levels = 1:2,
                       labels = c("gini", "entropy")),
    n_estimators = as.numeric(n_estimators),
    min_samples_split = as.numeric(min_samples_split),
    min_samples_leaf = as.numeric(min_samples_leaf)
  ) %>%
  mutate(across(c(max_depth, max_features, criterion), as.character))


rf_stdb <- train
```

```{r}
# add any metric columns you actually have in train_tbl
candidate_metrics <- c(
  "cv_ap", "cv_auc",
  "cv_mcc", "cv_f1", "cv_bal_acc",
  "cv_recall", "cv_specificity"
)

# pretty labels for table
label_map <- c(
  cv_ap       = "Mean AP",
  cv_auc      = "Mean AUROC",
  cv_mcc      = "Mean MCC",
  cv_f1       = "Mean F1",
  cv_bal_acc   = "Balanced Accuracy",
  cv_recall  = "Sensitivity",
  cv_specificity = "Specificity"
)

metrics_present <- intersect(candidate_metrics, names(train_tbl))

RF_summary <- train_tbl %>%
  select(run, all_of(metrics_present)) %>%
  pivot_longer(cols = -run, names_to = "metric", values_to = "value") %>%
  group_by(metric) %>%
  summarize(
    Mean  = mean(value),
    SD    = sd(value),
    SE    = SD / sqrt(n()),
    t_val = qt(0.975, df = n() - 1),
    `CI Low`  = Mean - t_val * SE,
    `CI High` = Mean + t_val * SE,
    Min   = min(value),
    Max   = max(value),
    .groups = "drop"
  ) %>%
  mutate(Metric = label_map[metric]) %>%
  select(Metric, Mean, SD, SE, `CI Low`, `CI High`, Min, Max)

RF_summary_gt <- metric_summary %>%
  gt() %>%
  fmt_number(
    columns = c(Mean, SD, SE, `CI Low`, `CI High`, Min, Max),
    decimals = 3
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  # ✅ Add spacing for readability
  cols_width(
    Metric ~ px(90),          # wider first column
    everything() ~ px(90)      # widen all numeric columns
  ) %>%
  tab_options(
    #column_label_padding = px(10),  # space above/below column headers
    data_row.padding = px(6)        # space in the cell rows
  )

RF_summary_gt

gtsave(RF_summary_gt, "final_plots/RF_train_stats.html")
```

```{r}
# Bind and create row groups
combined_summary <- bind_rows(
  RF_summary %>% mutate(Model = "Random Forest"),
  GB_summary %>% mutate(Model = "Gradient Boosting")
) %>%
  select(Model, Metric, Mean, SD, SE, `CI Low`, `CI High`, Min, Max) %>%
  arrange(factor(Model, levels = c("Random Forest", "Gradient Boosting")),
          Metric)

# Build stratified gt table (row groups = models)
combined_summary_gt <- combined_summary %>%
  gt(groupname_col = "Model") %>%
  fmt_number(
    columns = c(Mean, SD, SE, `CI Low`, `CI High`, Min, Max),
    decimals = 3
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),          # <-- bold the row group headers
    locations = cells_row_groups()
  ) %>%
  cols_width(
    Metric ~ px(160),
    everything() ~ px(90)
  )

combined_summary_gt

gtsave(combined_summary_gt, "final_plots/combined_train_stats.html")
```

```{r}
train_scores <- train |> select(run, cv_ap)

train_scores_unique <- train_scores %>%
  distinct(run, .keep_all = TRUE)

train <- train %>%
  select(-cv_ap, -run) %>%
  mutate(
    across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(train, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# Summarize F1 stats per run
train_summary <- train_scores %>%
  group_by(run) %>%
  summarise(
    mean_f1 = mean(cv_ap, na.rm = TRUE),
    min_f1 = min(cv_ap, na.rm = TRUE),
    max_f1 = max(cv_ap, na.rm = TRUE)
  )

plot_f1 <- ggplot(train_scores_unique, aes(x = factor(run), y = cv_ap)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "Mean Average Precision", title = "Mean Average Precision") +
  ylim(0, 0.8)

(param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )

ggsave("final_plots/RF_metrics.png", plot = last_plot(),
       width = 10, height = 8, units = "in", dpi = 300)
```

## Test Summary

```{r}
cc <- c(max_features = "character")  # add more like criterion="character" if needed

read_one <- function(path) {
  read.csv(path, header = TRUE, stringsAsFactors = FALSE, colClasses = cc)
}

test1 <- read_one('final_plots/RF1_Top5_Test.csv')
test2 <- read_one('../tuning2/final_plots/RF2_Top5_Test.csv')
test3 <- read_one('../tuning3/final_plots/RF3_Top5_Test.csv')
test4 <- read_one('../tuning4/final_plots/RF4_Top5_Test.csv')
test5 <- read_one('../tuning5/final_plots/RF5_Top5_Test.csv')

test1$run <- 'run1'
test2$run <- 'run2'
test3$run <- 'run3'
test4$run <- 'run4'
test5$run <- 'run5'

test <- dplyr::bind_rows(test1, test2, test3, test4, test5)

test_scores <- test |> select(run, test_avg_precision)

test_scores_unique <- test_scores %>%
  distinct(run, .keep_all = TRUE)

test <- test %>%
  select(-test_avg_precision, -run) %>%
  mutate(
    across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(test, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

plot_f1 <- ggplot(test_scores_unique, aes(x = factor(run), y = test_avg_precision)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "MCC Score", title = "MCC Score") +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Test - Top Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Validation Summary

```{r}
cc <- c(max_features = "character", max_depth = "character")  # add more like criterion="character" if needed

read_one <- function(path) {
  read.csv(path, header = TRUE, stringsAsFactors = FALSE, colClasses = cc)
}

val1 <- read_one('final_plots/RF1_Top5_Validation.csv')
val2 <- read_one('../tuning2/final_plots/RF2_Top5_Validation.csv')
val3 <- read_one('../tuning3/final_plots/RF3_Top5_Validation.csv')
val4 <- read_one('../tuning4/final_plots/RF4_Top5_Validation.csv')
val5 <- read_one('../tuning5/final_plots/RF5_Top5_Validation.csv')

val1$run <- 'run1'
val2$run <- 'run2'
val3$run <- 'run3'
val4$run <- 'run4'
val5$run <- 'run5'

val <- dplyr::bind_rows(val1, val2, val3, val4, val5)

val_scores <- val |> select(run, val_avg_precision)

val_scores_unique <- val_scores %>%
  distinct(run, .keep_all = TRUE)

val <- val %>%
  select(-val_avg_precision, -run) %>%
  mutate(
    across(everything(), as.factor))
```

```{r}
#| fig-height: 8
#| fig-width: 8

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(val, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

plot_f1 <- ggplot(val_scores_unique, aes(x = factor(run), y = val_avg_precision)) +
  geom_col(fill = "steelblue", color = 'black') +
  theme_minimal() +
  labs(x = "Model Run", y = "MCC Score", title = "MCC Score") +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Validation - Top Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## All Datasets Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(500, 1000)),
    max_depth = factor(max_depth, levels = c(10, 12, 16, 20, "None")),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.3", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(5, 10, 20, 50)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: avg_precision Score Plot ---
# Bring in CV, Test, and Val avg_precision side-by-side by run
RF1_scores <- list(
#  train_scores %>% select(run, cv_avg_precision),
  test_scores  %>% select(run, test_avg_precision),
  val_scores   %>% select(run, val_avg_precision)
) %>%
  purrr::reduce(left_join, by = "run")

# Long format: set = cv_avg_precision / test_avg_precision / val_avg_precision
scores_long <- RF1_scores %>%
  pivot_longer(
    cols = -run,
    names_to = "set",
    values_to = "ap"
  )

# One value per run & set (mean is trivial if already one row/run)
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(mean_ap = mean(ap, na.rm = TRUE), .groups = "drop")

# Ensure legend order matches your palette
summary_scores <- summary_scores %>%
  mutate(set = factor(set, levels = c("test_avg_precision", "val_avg_precision")))

plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_ap, fill = set)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6, color = "black") +
  geom_text(
    aes(y = mean_ap + 0.04, label = sprintf("%.2f", mean_ap), group = set),
    position = position_dodge(width = 0.7), size = 3, vjust = 0
  ) +
  scale_fill_manual(
    values = c("test_avg_precision" = "#2ca02c", "val_avg_precision" = "#ff7f0e"),
    name   = "Score Type",
    labels = c("Test", "Evaluation")
  ) +
  theme_minimal() +
  labs(x = "Model Run", y = "Average Precision") +
  coord_cartesian(ylim = c(0, 0.7))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1))

final_plot


```

## Upset Plot

```{r}
# Step 1: Filter and add a unique row ID for each model
filtered <- train %>%
  select(
    n_estimators,
    max_depth,
    min_samples_split,
    min_samples_leaf,
    criterion,
    max_features
  ) %>%
  # filter(
  #   criterion == "entropy",
  #   max_features == 0.5
  # ) %>%
  #select(-criterion, -max_features) %>%
  mutate(row_id = row_number())  # Add unique row ID

# Step 2: Convert to long format and flag presence
filtered_long <- filtered %>%
  pivot_longer(cols = -row_id, names_to = "param", values_to = "value") %>%
  mutate(present = TRUE) %>%
  unite("feature", param, value, sep = "_")

# Step 3: Convert to binary wide format
binary_data <- filtered_long %>%
  pivot_wider(
    id_cols = row_id,
    names_from = feature,
    values_from = present,
    values_fill = list(present = FALSE)
  ) %>%
  select(-row_id)

# Step 4: Make the upset plot
upset(
  data = binary_data,
  intersect = colnames(binary_data),
  name = "Hyperparameter Combination",
  # remove the numbers above the top bars
  base_annotations = list(
    "Intersection size" = intersection_size(counts = FALSE)
  ),
  # make the matrix dots/lines smaller via the matrix arg
  matrix = intersection_matrix(
    geom    = geom_point(size = 0.75),          # smaller nodes
    segment = geom_segment(linewidth = 0.2)     # thinner connecting lines
  ),
  height_ratio = 1.3,
  width_ratio  = 0.25,
  # optional: tidy up set-size labels
  set_sizes = upset_set_size() +
    theme(axis.text.x = element_text(hjust = 1, vjust = 1, size = 8))
)

ggsave("final_plots/RF_upset.png", plot = last_plot(),
       width = 12, height = 8, units = "in", dpi = 300)
```

## Validation and Test Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

# Convert to factors with correct level ordering
all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(50, 100, 500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(3, 7, 12, 20, 50, 100)),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 50, 100, 200, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 5, 10, 20, 50))
  )


# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: MCC Score Plot ---
RF1_scores <- test_scores_unique %>%
  left_join(val_scores_unique, by = "run")

scores_long <- RF1_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean MCC Score
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(mean_f1 = mean(f1, na.rm = TRUE), .groups = "drop")

# Plot with text labels for mean
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    color = 'black',
    stat = "identity"
  ) +
  geom_text(
    data = summary_scores,
    aes(
      x = factor(run),
      y = mean_f1 + 0.04,
      label = sprintf("%.2f", mean_f1),
      group = set
    ),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_ap" = "#1f77b4", "test_avg_precision" = "#2ca02c", "val_avg_precision" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "MCC Score",
    title = "MCC Score (Mean Highlighted)",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "Test and Validation Datasets – Top Models",
  )

final_plot
```

# Top 5 Models

## Training Summary

```{r}
# train <- read.csv('RF_summaries/RF1_Top5_Training.csv', header=TRUE)
# 
# train_scores <- train |> select(run, cv_ap)
# 
# # train_scores_unique <- train_scores %>%
# #   distinct(run, .keep_all = TRUE)
# 
# train <- train %>%
#   select(-cv_ap, -run) %>%
#   mutate(
#     # Recode first, then convert to factor
#     max_features = case_when(
#       max_features == "1" ~ "0.5",
#       max_features == "2" ~ "sqrt",
#       max_features == "3" ~ "log2",
#       TRUE ~ as.character(max_features)  # <-- required fallback
#     ),
#     criterion = case_when(
#       criterion == "1" ~ "gini",
#       criterion == "2" ~ "entropy",
#       TRUE ~ as.character(criterion)
#     )
#   ) %>%
#   mutate(across(everything(), as.factor)) 
```

```{r}
# #| fig-height: 8
# #| fig-width: 8
# 
# train_scores_grouped <- train_scores %>%
#   group_by(run) %>%
#   mutate(instance = row_number()) %>%
#   ungroup()
# 
# param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")
# 
# plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")
# 
# param_plots <- map2(param_cols, plot_colors, function(param, color) {
#   ggplot(train, aes_string(x = param)) +
#     geom_bar(fill = color, color = "black") +
#     theme_minimal() +
#     labs(title = param, x = NULL, y = "Count")
# })
# 
# param_plot_grid <- wrap_plots(param_plots, ncol = 3)
# 
# train_summary <- train_scores_grouped %>%
#   group_by(run) %>%
#   summarise(
#     min_f1 = min(cv_ap, na.rm = TRUE),
#     max_f1 = max(cv_ap, na.rm = TRUE),
#     mean_f1 = mean(cv_ap, na.rm = TRUE),
#     .groups = "drop"
#   )
# 
# plot_f1 <- ggplot(train_summary, aes(x = factor(run), y = mean_f1)) +
#   geom_col(fill = "steelblue", width = 0.6) +
#   geom_errorbar(
#     aes(ymin = min_f1, ymax = max_f1),
#     width = 0.2,
#     color = "black",
#     size = 0.6
#   ) +
#   geom_text(
#     aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
#     size = 3.5
#   ) +
#   theme_minimal() +
#   labs(
#     x = "Model Run",
#     y = "MCC Score",
#     title = "MCC Score per Run (Mean with Range)"
#   ) +
#   ylim(0, 0.75)
# 
# (param_plot_grid / plot_f1) + 
#   plot_layout(heights = c(3, 2)) +
#   plot_annotation(
#     title = "Training - Top 5 Models",
#     theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
#   )
```

## Test Summary

```{r}
test <- read.csv('RF_summaries/RF1_Top5_Test.csv', header=TRUE)

test <- test %>%
  mutate(max_depth = replace_na(as.character(max_depth), "None"))

test_scores <- test |> select(run, test_avg_precision)

test_scores_unique <- test_scores %>%
  distinct(run, .keep_all = TRUE)

test <- test %>%
  select(-test_avg_precision, -run) %>%
  mutate(
    # Recode first, then convert to factor
    max_features = case_when(
      max_features == "1" ~ "0.5",
      max_features == "2" ~ "sqrt",
      max_features == "3" ~ "log2",
      TRUE ~ as.character(max_features)  # <-- required fallback
    ),
    criterion = case_when(
      criterion == "1" ~ "gini",
      criterion == "2" ~ "entropy",
      TRUE ~ as.character(criterion)
    )
  ) %>%
  mutate(across(everything(), as.factor)) 
```

```{r}
#| fig-height: 8
#| fig-width: 8

test_scores_grouped <- test_scores %>%
  group_by(run) %>%
  mutate(instance = row_number()) %>%
  ungroup()

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(test, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

test_summary <- test_scores_grouped %>%
  group_by(run) %>%
  summarise(
    min_f1 = min(test_avg_precision, na.rm = TRUE),
    max_f1 = max(test_avg_precision, na.rm = TRUE),
    mean_f1 = mean(test_avg_precision, na.rm = TRUE),
    .groups = "drop"
  )

plot_f1 <- ggplot(test_summary, aes(x = factor(run), y = mean_f1)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_errorbar(
    aes(ymin = min_f1, ymax = max_f1),
    width = 0.2,
    color = "black",
    size = 0.6
  ) +
  geom_text(
    aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
    size = 3.5
  ) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "MCC Score",
    title = "MCC Score (Mean with Range)"
  ) +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## Validation Summary

```{r}
val <- read.csv('RF_summaries/RF1_Top5_Validation.csv', header=TRUE)

val <- val %>%
  mutate(max_depth = replace_na(as.character(max_depth), "None"))

val_scores <- val |> select(run, val_avg_precision)

val_scores_unique <- val_scores %>%
  distinct(run, .keep_all = TRUE)

val <- val %>%
  select(-val_avg_precision, -run) %>%
  mutate(
    # Recode first, then convert to factor
    max_features = case_when(
      max_features == "1" ~ "0.5",
      max_features == "2" ~ "sqrt",
      max_features == "3" ~ "log2",
      TRUE ~ as.character(max_features)  # <-- required fallback
    ),
    criterion = case_when(
      criterion == "1" ~ "gini",
      criterion == "2" ~ "entropy",
      TRUE ~ as.character(criterion)
    )
  ) %>%
  mutate(across(everything(), as.factor)) 
```

```{r}
#| fig-height: 8
#| fig-width: 8

val_scores_grouped <- val_scores %>%
  group_by(run) %>%
  mutate(instance = row_number()) %>%
  ungroup()

param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(val, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

val_summary <- val_scores_grouped %>%
  group_by(run) %>%
  summarise(
    min_f1 = min(val_avg_precision, na.rm = TRUE),
    max_f1 = max(val_avg_precision, na.rm = TRUE),
    mean_f1 = mean(val_avg_precision, na.rm = TRUE),
    .groups = "drop"
  )

plot_f1 <- ggplot(val_summary, aes(x = factor(run), y = mean_f1)) +
  geom_col(fill = "steelblue", width = 0.6) +
  geom_errorbar(
    aes(ymin = min_f1, ymax = max_f1),
    width = 0.2,
    color = "black",
    size = 0.6
  ) +
  geom_text(
    aes(label = sprintf("%.2f", mean_f1), y = mean_f1 + 0.06),
    size = 3.5
  ) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "MCC Score",
    title = "MCC Score per Run (Mean with Range)"
  ) +
  ylim(0, 0.75)

(param_plot_grid / plot_f1) + 
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "Validation - Top 5 Models",
    theme = theme(plot.title = element_text(size = 18, face = "bold", hjust = 0.5))
  )
```

## All Datasets Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

all <- all %>%
  mutate(max_depth = replace_na(max_depth, "None")) |>
  mutate(
    n_estimators = factor(n_estimators, levels = c(500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(100, 500, "None")),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 100, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 10, 50))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

ggsave(
  filename = "plots/RF_hyperparam_summary.png",       # File name (can be .png, .pdf, .jpeg, etc.)
  plot = param_plot_grid,          # Or specify a plot object, e.g., plot = p
  device = NULL,               # Auto-detects from file extension, or specify like "png", "pdf"
  path = NULL,                 # Optional: directory to save file in
  scale = 1,                   # Scaling factor
  width = 6,                   # Width in units
  height = 6,                  # Height in units
  units = "in",                # Units: "in", "cm", "mm"
  dpi = 300,                   # Dots per inch (resolution)
  limitsize = TRUE             # If TRUE, limits size to 50x50 inches
)
```

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- read_csv('final_plots/RF_top_models.csv')

all <- all %>%
  mutate(max_depth = replace_na(max_depth, "None")) |>
  mutate(
    n_estimators = factor(n_estimators, levels = c(500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(100, 500, "None")),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 100, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 10, 50))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count")
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

param_plot_grid

ggsave(
  filename = "final_plots/RF_hyperparam_summary.png",       # File name (can be .png, .pdf, .jpeg, etc.)
  plot = param_plot_grid,          # Or specify a plot object, e.g., plot = p
  device = NULL,               # Auto-detects from file extension, or specify like "png", "pdf"
  path = NULL,                 # Optional: directory to save file in
  scale = 1,                   # Scaling factor
  width = 6,                   # Width in units
  height = 6,                  # Height in units
  units = "in",                # Units: "in", "cm", "mm"
  dpi = 300,                   # Dots per inch (resolution)
  limitsize = TRUE             # If TRUE, limits size to 50x50 inches
)
```

## Validation and Test Summary

```{r}
#| fig-height: 8
#| fig-width: 8

# Combine all
all <- bind_rows(list(test, val))

all <- all %>%
  mutate(
    n_estimators = factor(n_estimators, levels = c(500, 1000, 1500)),
    max_depth = factor(max_depth, levels = c(100, 500, "None")),
    max_features = factor(max_features, levels = c("sqrt", "log2", "0.5")),
    criterion = factor(criterion, levels = c("gini", "entropy")),
    min_samples_split = factor(min_samples_split, levels = c(20, 100, 500)),
    min_samples_leaf = factor(min_samples_leaf, levels = c(1, 10, 50))
  )

# Color palette
param_cols <- c("n_estimators", "max_depth", "min_samples_split", "min_samples_leaf", "max_features", "criterion")

plot_colors <- brewer.pal(n = length(param_cols), name = "Dark2")

param_plots <- map2(param_cols, plot_colors, function(param, color) {
  ggplot(all, aes_string(x = param)) +
    geom_bar(fill = color, color = "black") +
    theme_minimal() +
    labs(title = param, x = NULL, y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
})

param_plot_grid <- wrap_plots(param_plots, ncol = 3)

# --- Step 4: MCC Score Plot ---
RF1_scores <- test_scores %>%
  left_join(val_scores, by = "run")

scores_long <- RF1_scores %>%
  pivot_longer(cols = -run, names_to = "set", values_to = "f1")

# Summarize mean, min, and max MCC Score
summary_scores <- scores_long %>%
  group_by(run, set) %>%
  summarise(
    mean_f1 = mean(f1, na.rm = TRUE),
    ymin = min(f1, na.rm = TRUE),
    ymax = max(f1, na.rm = TRUE),
    .groups = "drop"
  )

# Final plot with bars, error bars, and mean labels
plot_f1 <- ggplot(summary_scores, aes(x = factor(run), y = mean_f1, fill = set)) +
  geom_col(
    position = position_dodge(width = 0.7),
    width = 0.6,
    # color = 'black',
    stat = "identity"
  ) +
  geom_errorbar(
    data = summary_scores,
    aes(x = factor(run), ymin = ymin, ymax = ymax, group = set),
    position = position_dodge(width = 0.7),
    width = 0.25,
    size = 0.75,
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text(
    data = summary_scores,
    aes(x = factor(run), y = ymax + 0.06, label = sprintf("%.2f", mean_f1), group = set),
    position = position_dodge(width = 0.7),
    size = 3,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = c("cv_ap" = "#1f77b4", "test_avg_precision" = "#2ca02c", "val_avg_precision" = "#ff7f0e")) +
  theme_minimal() +
  labs(
    x = "Model Run",
    y = "MCC Score",
    title = "Mean MCC Score and Range",
    fill = "Score Type"
  ) +
  coord_cartesian(ylim = c(0, 0.75))

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

# --- Step 5: Combine all plots with title ---
final_plot <- (param_plot_grid / plot_f1) +
  plot_layout(heights = c(3, 1)) +
  plot_annotation(
    title = "All Datasets – Top Models",
  )

final_plot
```

## Upset Plot

```{r}
#| fig-height: 8
#| fig-width: 8

# Step 1: Filter and add a unique row ID for each model
filtered <- all %>%
  mutate(max_depth = replace_na(max_depth, "None")) |>
  select(
    n_estimators,
    max_depth,
    min_samples_split,
    min_samples_leaf,
    criterion,
    max_features
  ) %>%
  mutate(row_id = row_number())  # Add unique row ID

# Step 2: Convert to long format and flag presence
filtered_long <- filtered %>%
  pivot_longer(cols = -row_id, names_to = "param", values_to = "value") %>%
  mutate(present = TRUE) %>%
  unite("feature", param, value, sep = "_")

# Step 3: Convert to binary wide format
binary_data <- filtered_long %>%
  pivot_wider(
    id_cols = row_id,
    names_from = feature,
    values_from = present,
    values_fill = list(present = FALSE)
  ) %>%
  select(-row_id)

# Step 4: Make the upset plot
upset(
  binary_data,
  intersect = colnames(binary_data),
  name = "Hyperparameter Combination",
  height_ratio = 1.3
)

ggsave(
  filename = "final_plots/RF_upset.png",       # File name (can be .png, .pdf, .jpeg, etc.)
  plot = last_plot(),          # Or specify a plot object, e.g., plot = p
  device = NULL,               # Auto-detects from file extension, or specify like "png", "pdf"
  path = NULL,                 # Optional: directory to save file in
  scale = 1,                   # Scaling factor
  width = 6,                   # Width in units
  height = 6,                  # Height in units
  units = "in",                # Units: "in", "cm", "mm"
  dpi = 300,                   # Dots per inch (resolution)
  limitsize = TRUE             # If TRUE, limits size to 50x50 inches
)
```

```{r}
#| fig-height: 8
#| fig-width: 8

# Step 1: Filter and add a unique row ID for each model
filtered <- all %>%
  select(
    n_estimators,
    max_depth,
    min_samples_split,
    min_samples_leaf,
    criterion,
    max_features
  ) %>%
  filter(
    criterion == "entropy",
    max_features == 0.5
  ) %>%
  select(-criterion, -max_features) %>%
  mutate(row_id = row_number())  # Add unique row ID

# Step 2: Convert to long format and flag presence
filtered_long <- filtered %>%
  pivot_longer(cols = -row_id, names_to = "param", values_to = "value") %>%
  mutate(present = TRUE) %>%
  unite("feature", param, value, sep = "_")

# Step 3: Convert to binary wide format
binary_data <- filtered_long %>%
  pivot_wider(
    id_cols = row_id,
    names_from = feature,
    values_from = present,
    values_fill = list(present = FALSE)
  ) %>%
  select(-row_id)

# Step 4: Make the upset plot
upset(
  binary_data,
  intersect = colnames(binary_data),
  name = "Hyperparameter Combination",
  height_ratio = 1.3,
  min_size = 2
)
```

I chose the right-most combination because it contains more hyperparameter values that were chosen more often by good models.
